---
title: "Biol 3872 Lab8-Assignment"
output: html_document
date: "`r Sys.Date()`"
author: ""
student_id: ""
---

\
**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Arun Oakley-Cogan") and Banner ID ("B00123456") in `student_id`.\
2. Install and/or load the packages required to run your code chunks.

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

```{r}
# LOAD NECCESSARY PACKAGES HERE (as you go through questions)
```

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 1 {3 points}

A)  Load in all the data for this assignment. Start with the `Groundfish_surveys.csv` file as an `sf` object called `trawls` from the lab 7 folder. Set the coordinate reference system (CRS) to `EPSG:4326` (WGS84, unprojected). Load in `ne_countries` at `medium` scale from `rnaturalearth`, call this `coastline`. Lastly, load four `SpatRaster` layers into an object called `marine_stack` from `sdmpredictors`:

-   `BO2_tempmean_bdmax` (Sea water temperature average at seafloor)
-   `BO2_nitratemean_bdmax` (Nitrate concentration average at seafloor)
-   `BO21_phosphatemean_bdmax` (Phosphate concentration average at seafloor)
-   `BO21_dissoxmean_bdmax` (Dissolved oxygen concentration average at seafloor)

B)  Convert the `trawls` dataframe to a `SpatVector` called `trawls_vect`. Use these vector points to `extract` values from all raster layers in `marine_stack` into a `data.frame` called `env`. Bind `env` to `trawls_vect` and convert this to an `sf` object called `trawls_env`.

C)  Use the extent (bounding box) of `trawls_env` to `mask` and `crop` all `marine_stack` layers. Call this `marine_stack_NW`. Checked that this worked by plotting.
```{r}

# YOUR CODE HERE

# A)


# B)


# C)


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 2 {2 points}

A)  Interpolate the `marine_stack_NW` `SpatRaster` layers using bilinear averaging and **increase** the resolution (more cells) by a factor of 5, call this `marine_stack_NW_fine`.

B)  Project `marine_stack_NW_fine` to a Prince Edward Island Stereographic projection (`EPSG:2954`) and call this `marine_stack_NW_fine_proj`.
```{r}
# YOUR CODE HERE

# A)


# B)

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 3 {3 points)

A.  Filter the coastline object for only the Canadian polygon, store in an object called `canada`.

B.  Project `canada` into an object called `canada_proj` using the Prince Edward Island Stereographic projection. Calculate the area in $km^2$ of both `canada` and `canada_proj` and store in two separate objects called `area_spherical` and `area_projected`.

C.  Calculate and store the difference between `area_spherical` and `area_projected` in an object called `coastline_diff`.
```{r}
# YOUR CODE HERE

# A)


#B)


#C)


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 4 {3 points}

A. Using `tidyterra`, create an object called `marine_stack_NW_O2` which selects the dissolved oxygen layer from `marine_stack_NW`, and `mutate` on a new layer called `O2_ppm` which calculates dissolved oxygen in ppm by multiplying by 0.032.

B. Convert `trawls_env` to a spatial vector (`SpatVector`) called `trawls_env_vect`, filter it to include only longhorn sculpin (*Myoxocephalus octodecemspinosus*), and `mutate` a new column called `log_cpue` which calculates the natural log of catch per unit effort.

C. Create a ggplot called `tidyterra_plot` using `geom_spatvector` and `geom_spatraster` which maps dissolved oxygen in ppm from `marine_stack_NW_O2` using the `mako` viridis palette, overlaid with the points in `trawls_env_vect`. Color the points by temperature using the `magma` viridis palette, size them by `log_cpue`, and set them to 50% transparency. Think about whether each of these elements should be inside or outside of `aes()`. Name the relevant legends 'Oxygen (ppm)', 'Temperature (°C)', and 'log CPUE'.
```{r}
# YOUR CODE HERE

# A)


# B)


# C)


```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 5 {4 points}

A)  Store the bounding box of `trawls` in a variable called `ext_unproj` (this should be a `bbox` class, not a `SpatExtent`).

B)  Project `trawls` to `EPSG:2954` (Prince Edward Island Stereographic / NAD83) and store its `bbox` in a variable called `ext_proj`. Call the projected trawl data `trawls_proj`.

C)  Construct grids for `trawls` and `trawls_proj` with cell sizes of 0.25 degrees and 20 km^2^, respectively. Call these grids `grid_unproj` / `grid_proj`, respectively.

D)  Fix the **explicit** `if`/`else` statement which tests which grid version has a finer representation (more grid cells) of our NW Atlantic region, and **stores** the result to a variable called `answer`. The correct version should effectively concatenate outputs as a `character` string, storing a variable (`answer`) stating which version achieves finer resolution by calculating the difference in the number of grid cells.\
```{r}
# YOUR CODE HERE

# A)


# B)


# C)


# D)
# CODE TO FIX
#if {ncell(grid_unproj) >> ncell(grid_proj)}{
#  answer <- paste0("The geographic version has ", as.numeric(ncell(grid_unproj) - #count(grid_proj)), " more grid cells")
#} else {
#  paste0("The projected version has ", as.charcter(ncell(grid_proj) - #ncell(grid_unproj)), " more grid cells")
#}

```
:::

\

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"}

## Task 6 {2 points}

A)  Change the name of each layer of `marine_stack_NW_fine_proj` to `b_temp`, `b_nitrate`, `b_phosphate`, `b_dissox`, making sure to assign these simpler names to the correct layer.

B)  Write each layer of `marine_stack_NW_fine_proj` to a `GeoTIFF` file. Use `paste0` to concatenate strings in the `filenames` argument of `writeRaster`: These strings should direct outputs to the `Data` folder, name each output file from the `marine_stack_NW_fine_proj` layer name, and end in a `_shelf` suffix (e.g., `layer1_shelf.tif`).
```{r}
# YOUR CODE HERE
# A)


# B)

```
:::

\
